---
title: "Data Wrangling Class Notes"
author: "Erich Denk"
date: "February 4, 2019"
output: pdf_document
---

#Wrangling
90% of data science is data wrangling; including manipulating, piping, joining, reshaping.

##What is Wrangling?
The process of transforming from one format to another. 
Converting the structure to facilitate some analysis
  -altering the unit of analysis
  -changing from a wide to a long structure
  -summarizing data across subgroups

Data rarely works out of the box. Sometimes contructed for a different purpose. Requires adding additional information.

Codebooks offer clues but not the story. Pay attention to what is actually there.

When you think about Data Generation, consider `data provenance` and `automated data`. Who is touching the data, don't ever assume that an automated process was foolproof..

Almost everything you need to use is in the tidyverse.
`readr` for reading data in
`tibble` for tidy data structures
`dplyr` for data manipulation
`ggplot2` for visualization
`tidyr` for cleaning
`purrr` for funtional programming toolkit

```{r}
library(tidyverse)

```

##Tidy Data

The philosophy of the tidyverse. Basically, data can be presented in a lot of different ways that say the same thing. But really we want things in our basic row column format. For things to be tidy:
  1. Each variable must have its own column
  2. Each observation must have its own row
  3. Each value has its own cell

Tibble() data frames. They are data frames but they tweak some older behaviors to make life easier. They just have better features...

The main differences between the `tibbles` and `data.frames`.
  1. Tibbles have a refined print method that shows only the first 1 rows and all columns that fit on the screen.
  2. Tibbles are strict. they throw lots of errors that we can catch early. 


##Manipulating

The `dplyr` package offers an intuitive verb based approach to data management in R.
  - grammar of data manipulation or data visualization
  - all about intuitive naming convention to data management tasks

There are 6 main `dplyr verbs`
- select, filter, arrange, mutate, summarise, groupby

They all work similarly.
  1. The first argument is a data frame
  2. The subsequent arguments say what to do with the data frame
  3. Result is new data frame
  
###Select
Lets take a look using some built in data
```{r}
?presidential
dat <- presidential
head(dat)
```

Say we just want the name and the party of the presidents
```{r}
select(dat, name, party)
```

We can also get all variables in between name and end.  Like a range of variables
```{r}
select(dat, name:end)
```
Note that we don't need to put quotations around the variable names. Which we can do 
The order by which you specify data is helpful in organizing you can also rename simultaneously.

```{r}
select(dat, president = name, 
       startdate = start,
       enddate = end)
```

You can also use negate - to drop the variables you don't want...

###Filter
We can use boolean concepts to figure things out and filter to just data we want to see or not to see. 
```{r}
filter(dat, party == "Republican")
```

###Arrange
Does what you think it would . You can put things in the order that you want.  STATA does it too. 
```{r}
arrange(dat, party)

arrange(dat, desc(start))

```

###Mutate
A key use. Make new variables. Cold War official start date variable.
```{r}
dat <- mutate(dat, CW = start <='1990-03-11',
       CW = as.numeric(CW))

#Or you can also only see the variables that you created.
transmute(dat, CW = start <= '1990-03-11',
          CW = as.numeric(CW))
```

###Summarize Data
```{r}
summarize(presidential,
          days_in_office = mean(end-start),
          max = max(end-start),
          min = min(end-start))
```

###Grouping by Variable Characteristics
group_by is incredibly helpful. you can have "good" and "bad" groups. 
So we have some grouping variable we want to group by. Under the hood it will create a way to separate them.
```{r}
#group by party
x <- group_by(dat, party)
summarize(x,min_in_office = min(end-start))
```

###Other functions in the tidyverse
tally() or count() offers quick counts of variables which can be helpful

recode() works inside mutate to create a new variable. Say switch to 0,1 binary 
```{r}
mutate(dat, party = recode(party, 'Republican' = 1, 'Democratic'=0))
```
And so much more!
sample_n() grab an N random sample of your data
sample_frac() grab a random fraction
top_n get the top N number of entries
slice() grab sepcific row ranges
glimpse() quick preview of the data

##Piping
Think about how we could combine `dplyr` functions. So we might want to do a series of things
```{r}
x <- filter(presidential, party=='Republican')
x <-  group_by(x,name)
x <- transmute(x,t_in_office = end- start)
x <- arrange(x,t_in_office)
```
Or we could nest everything together in nests, but you have to read inside out. Instead we can use the pipe. Data goes in and new comes out

```{r}
dat%>%
  filter(party=='Republican') %>%
  group_by(name) %>%
  transmute(t_in_office = end-start) %>%
  arrange(t_in_office)
```

Two things to keep in mind when piping ctrl + shift + m
  1. Functions must be linked with %>%
  2. When functions have multiple arguments, point to where the data should go with a period (.)
  
##Joining
This provides an SQL framework that is easier to read and more efficient.
_join is the unifying verb in a range of ways to combine them
  - left_join()
  - right_join() 
  - inner_join() keeps only those that match
  - full_join() good for keeping all the data and combining where it fits
  - anti_join() good for finding the unique observations that are specific to one dataset
Class presentation has a great visualization of these five types of joins. 

bind_rows() you want to stack them ontop of each other. Creates columns of NA for where it doesn't work for you. Also, can work for multiple datasets. 

bind_cols() doesn't have as much use. 

There's also set theory uses but they aren't that helpful

##Reshaping
tidyr is used to reshape data. It eases dropping missing values, filling missing values, and separating columns into rows. Like you did for your thesis.

The two main ones we will use include `gather()` and `spread()`
  - gather moves from wide to long
  - spread moves from long to wide
